Wind Veer and Speed in Turbulent Ekman Flow Part I: Scaling analysis and universal profile model

### 2024-11-07 write1/manuscript.tex
Minor edits by Hauke

Another common measure of scale separation is the Reynolds number $\RE_D = G D /\nu $  defined by the laminar Ekman layer thickness $D=\sqrt{2\nu/f}$.

@@

\section{Laminar Ekman solution with consideration of inner layer}
\label{app:ekman_solution}
The following Ekman system is obtained by averaging the Navier--Stokes equations horizontally and over time and neglecting the turbulent transport terms (turbulence can then be considered via the eddy-viscosity concept through variations in the viscosity $\nu$):
\begin{subequations}
\begin{eqnarray}
  \left(\begin{matrix}
    \p_t U\\
    \p_t V  
  \end{matrix}\right)&=&\left(\begin{matrix}
     fV &+& \nu \p_z^2 U\\ 
    -f(U-G) &+& \nu \p_z^2 V
  \end{matrix}\right)\\ 
  \Rightarrow \partial_t (U+iV) &=& f(V-i(U-G))) + \nu \p_z^2(U+iV)
\end{eqnarray}
In stationary conditions, this system is solved by
\begin{eqnarray}
  \hat{u}(z) &=& U_{\infty} + e^{-\gamma z} \left[A \cos\gamma z + B \sin\gamma z\right]   \\
  \hat{v}(z) &=& V_{\infty} + e^{-\gamma z} \left[-A \sin\gamma z + B \cos\gamma z\right]
\end{eqnarray} 
\end{subequations}
where the constants $U_\infty$, $V_\infty$ set the top boundary condition and $A$ and $B$ set the bottom boundary condition. 
%
The most common boundary condition for a surface Ekman layer is $A=U_{\infty}=G$, $B=0$, and $V_{\infty}=0$.
%
The lower boundary condition, however, neglects the existence of the surface layer, and it appears reasonable to define
$A=c G$ where $c<1$ is a constant that icorporates the increased shear in the surface layer.
%
Given a 'matching height' $z_{match}$ and normalized matching height $\xi=\gamma z_\text{match}$ in the upper part of the inner layer, we can match the Ekman profile
to the inner layer by letting
%
\begin{subequations}
\begin{eqnarray}
  \begin{matrix} 
    u(z_\text{match}) \equiv u_\text{match} &=& U_{\infty} + e^{-\gzm} \left [ A \cos\gzm + B\sin\gzm \right] \\
    v(z_\text{match}) \equiv v_\text{match} &=& V_{\infty} + e^{-\gzm} \left [-A \sin\gzm + B\cos\gzm \right]
  \end{matrix} 
\end{eqnarray}
\begin{eqnarray}
  \Rightarrow \left(\begin{matrix}
    u_\text{match}-U_{\infty} \\ 
    v_\text{match}-V_{\infty}
  \end{matrix}\right) = e^{-\gzm}\left(\begin{matrix}A \\ B \end{matrix}\right)\left(\begin{matrix}
    \cos\gzm &+ \sin\gzm \\ -\sin\gzm &+ \cos\gzm
  \end{matrix} \right) \\
\end{eqnarray}
\end{subequations}
Matching the profile at $\xi=0$, one obtains
$A = \Delta u_\text{match} $ and $B=-\Delta w_\text{match} $; and when the direction $Ox$ is aligned
with the geostrophic wind, we obtain the textbook-case $A=|\mathbf{G}|$ and $B=0$.

@@

\section{Matching the spanwise velocity profiles in the inner layer}
\label{app:log-coefficients}
The spanwise profile in vicinity of the surface is given by
$V/G=f_{V,\mathrm{visc}}\delta ^+$ with
\begin{subequations}
\begin{eqnarray}
  f_{V,\mathrm{visc}} &=& v_\mathrm{ref}\left(\omega_v z^+ -1 + e^{-\omega_v z^+}\right)\\
  f_{V,\mathrm{log}}   &=& a_\mathrm{log} + b_\mathrm{log} \log z^+ + c_\mathrm{log} z^+ 
\end{eqnarray}
\end{subequations}
Matching the profiles and gradient $z_0=10^+$ and the value at $z_1=0.3\delta^+$ yields
\begin{subequations} 
\begin{eqnarray}
  v_\mathrm{ref} \left(\omega_v z_{0} + e^{-\omega_v z_{0}}\right)
  & = v_{0}=& a_\mathrm{log} + b_\mathrm{log} \log z_{0} +c_\mathrm{log} z_{0}  \\
  &   v_{1}=& a_\mathrm{log} + b_\mathrm{log} \log z_{1} +c_\mathrm{log} z_{1} \\
  v_\mathrm{ref}\omega_z \left( 1 - e^{-\omega_z z_{0}}\right)
  & = d_{0}=& \frac{b_\mathrm{log}}{z_{10}} + c_\mathrm{log} 
\end{eqnarray}
\end{subequations} 
The gradient condition implies $b_\mathrm{log}=(d_{0} -c_\mathrm{log})z_{0}$, and yields
\begin{subequations}
\begin{eqnarray}
  %v_{0}&=& a_\mathrm{log} + z_{0}\left[ (d_0-c_\mathrm{log}) \log z_{0} +c_\mathrm{log}\right]\\
  v_0-z_0d_0\log z_0&=& a_\mathrm{log} + c_\mathrm{log} (z_0- z_0\log z_0) \\
  %v_1 &=& a_\mathrm{log} + z_0(d_0-c_\mathrm{log})\log z_1 +  c_\mathrm{log}z_1\\ 
  v_1-z_0d_0\log z_1 &=& a_\mathrm{log} +c_\mathrm{log} (z_1- z_0\log z_0)  \\
  \Rightarrow c_\mathrm{log}&=& \frac{ \Delta v - z_0d_0 \log z_1/z_0}{\Delta z} 
\end{eqnarray}
with $\Delta z=z_1-z_0$ and $\Delta v=v_1-v_0$.
%
Then, the coefficient $a_\mathrm{log}$ is estimated as 
\begin{eqnarray} 
  a_\mathrm{log} &=& v_0-z_0d_0\log z_0 -  \frac{ \Delta v - z_0d_0 \log z_1/z_0}{\Delta z}\left[
    z_0-z_0\log z_0 
    \right]. 
\end{eqnarray}
We note that $ \log(z_1/z_0) / (z_1-z_0) \rightarrow 0$ for large $z_1$, and as $z_1=0.3\delta^+$,
this implies that the second term in $c_\mathrm{log}$ only plays a role at low and intermediate $\RE$.
Then, $a_\mathrm{log}$ can be estimated as 
\begin{eqnarray}
  a_\mathrm{log} \simeq v_0 -z_0 \left[ d_0 \log  z_0 - \frac{\Delta v}{\Delta z} ( 1- \log z_0) \right] 
\end{eqnarray}
for large $\RE$. 
\end{subequations}

>>>>

Another common measure of scale separation is the Reynolds number 
  \RE_D = \frac{G D}{\nu} 
\end{equation}  defined by the laminar Ekman layer thickness $D=\sqrt{2\nu/f}$.

@@

\section*{Appendix A: Laminar Ekman solution with consideration of inner layer}
\label{app:ekman_solution}
The following Ekman system is obtained by averaging the Navier--Stokes equations horizontally and over time and neglecting the turbulent transport terms (turbulence can then be considered via the eddy-viscosity concept through variations in the viscosity $\nu$):
\begin{subequations}
\begin{eqnarray}
  \left(\begin{matrix}
    \p_t U\\
    \p_t V  
  \end{matrix}\right)&=&\left(\begin{matrix}
     fV &+& \nu \p_z^2 U\\ 
    -f(U-G) &+& \nu \p_z^2 V
  \end{matrix}\right)\\ 
  \Rightarrow \partial_t (U+iV) &=& f(V-i(U-G))) + \nu \p_z^2(U+iV)
\end{eqnarray}
In stationary conditions, this system is solved by
\begin{eqnarray}
  \hat{u}(z) &=& U_{\infty} + e^{-\gamma z} \left[A \cos\gamma z + B \sin\gamma z\right]   \\
  \hat{v}(z) &=& V_{\infty} + e^{-\gamma z} \left[-A \sin\gamma z + B \cos\gamma z\right]
\end{eqnarray} 
\end{subequations}
where the constants $U_\infty$, $V_\infty$ set the top boundary condition and $A$ and $B$ set the bottom boundary condition. 
%
The most common boundary condition for a surface Ekman layer is $A=U_{\infty}=G$, $B=0$, and $V_{\infty}=0$.
%
The lower boundary condition, however, neglects the existence of the surface layer, and it appears reasonable to define
$A=c G$ where $c<1$ is a constant that incorporates the increased shear in the surface layer.
%
Given a 'matching height' $z_{match}$ and normalized matching height $\xi=\gamma z_\text{match}$ in the upper part of the inner layer, we can match the Ekman profile
to the inner layer by letting
%
\begin{subequations}
\begin{eqnarray}
  \begin{matrix} 
    u(z_\text{match}) \equiv u_\text{match} &=& U_{\infty} + e^{-\gzm} \left [ A \cos\gzm + B\sin\gzm \right] \\
    v(z_\text{match}) \equiv v_\text{match} &=& V_{\infty} + e^{-\gzm} \left [-A \sin\gzm + B\cos\gzm \right]
  \end{matrix} 
\end{eqnarray}
\begin{eqnarray}
  \Rightarrow \left(\begin{matrix}
    u_\text{match}-U_{\infty} \\ 
    v_\text{match}-V_{\infty}
  \end{matrix}\right) = e^{-\gzm}\left(\begin{matrix}A \\ B \end{matrix}\right)\left(\begin{matrix}
    \cos\gzm &+ \sin\gzm \\ -\sin\gzm &+ \cos\gzm
  \end{matrix} \right) \\
\end{eqnarray}
\end{subequations}
Matching the profile at $\xi=0$, one obtains
$A = \Delta u_\text{match} $ and $B=-\Delta w_\text{match} $; and when the direction $Ox$ is aligned
with the geostrophic wind, we obtain the textbook-case $A=|\mathbf{G}|$ and $B=0$.

@@

\section*{Appendix B: Matching the spanwise velocity profiles in the inner layer}

\label{app:log-coefficients}
The spanwise profile in vicinity of the surface is given by
$V/G=f_{V,\mathrm{visc}}\delta ^+$ with
\begin{subequations}
\begin{eqnarray}
  f_{V,\mathrm{visc}} &=& v_\mathrm{ref}\left(\omega_v z^+ -1 + e^{-\omega_v z^+}\right)\\
  f_{V,\mathrm{log}}   &=& a_\mathrm{log} + b_\mathrm{log} \log z^+ + c_\mathrm{log} z^+ 
\end{eqnarray}
\end{subequations}
Matching the profiles and gradient $z_0=10^+$ and the value at $z_1=0.3\delta^+$ yields
\begin{subequations} 
\begin{eqnarray}
  v_\mathrm{ref} \left(\omega_v z_{0} + e^{-\omega_v z_{0}}\right)
  & = v_{0}=& a_\mathrm{log} + b_\mathrm{log} \log z_{0} +c_\mathrm{log} z_{0}  \\
  &   v_{1}=& a_\mathrm{log} + b_\mathrm{log} \log z_{1} +c_\mathrm{log} z_{1} \\
  v_\mathrm{ref}\omega_z \left( 1 - e^{-\omega_z z_{0}}\right)
  & = d_{0}=& \frac{b_\mathrm{log}}{z_{10}} + c_\mathrm{log} 
\end{eqnarray}
\end{subequations} 
The gradient condition implies $b_\mathrm{log}=(d_{0} -c_\mathrm{log})z_{0}$, and yields
\begin{subequations}
\begin{eqnarray}
  %v_{0}&=& a_\mathrm{log} + z_{0}\left[ (d_0-c_\mathrm{log}) \log z_{0} +c_\mathrm{log}\right]\\
  v_0-z_0d_0\log z_0&=& a_\mathrm{log} + c_\mathrm{log} (z_0- z_0\log z_0) \\
  %v_1 &=& a_\mathrm{log} + z_0(d_0-c_\mathrm{log})\log z_1 +  c_\mathrm{log}z_1\\ 
  v_1-z_0d_0\log z_1 &=& a_\mathrm{log} +c_\mathrm{log} (z_1- z_0\log z_0)  \\
  \Rightarrow c_\mathrm{log}&=& \frac{ \Delta v - z_0d_0 \log z_1/z_0}{\Delta z} 
\end{eqnarray}
with $\Delta z=z_1-z_0$ and $\Delta v=v_1-v_0$.
%
Then, the coefficient $a_\mathrm{log}$ is estimated as 
\begin{eqnarray} 
  a_\mathrm{log} &=& v_0-z_0d_0\log z_0 -  \frac{ \Delta v - z_0d_0 \log z_1/z_0}{\Delta z}\left[
    z_0-z_0\log z_0 
    \right]. 
\end{eqnarray}
We note that $ \log(z_1/z_0) / (z_1-z_0) \rightarrow 0$ for large $z_1$, and as $z_1=0.3\delta^+$,
this implies that the second term in $c_\mathrm{log}$ only plays a role at low and intermediate $\RE$.
Then, $a_\mathrm{log}$ can be estimated as 
\begin{eqnarray}
  a_\mathrm{log} \simeq v_0 -z_0 \left[ d_0 \log  z_0 - \frac{\Delta v}{\Delta z} ( 1- \log z_0) \right] 
\end{eqnarray}
for large $\RE$. 
\end{subequations}
